

%------------------------------------------------------------------------%



%---------------------------------------------------------------------------------Page 251 C-%

\noindent \textbf{Capacities of special channels}
\textbf{\emph{Lossless Channel}}\\\begin{itemize} \item For a lossless channel, the mutual information (information transfer) is equal to the input (source) entropy), and no source information is lost in transmission.\item It can be shown that $H(X|Y) = 0$ ( If $y_i$ is the output, there is certainty about the input). Also $I(X;Y) = H(X)$.
\item Consequently, the channel capacity per symbol is
\[ C_s = \mbox{ max }_{P(x_i)} H(X) = \mbox{log}_2m \]
where $m$ is the number of symbols in $X$.
\item For example, if there are $m=4$ input channels, then $C =  \mbox{log}_2 4 = 2$ b/symbol  \end{itemize}


%---------------------------------------------------------------------------------Page 251 D-%

\noindent \textbf{Capacities of special channels}
\textbf{\emph{Deterministic Channel}}:
\begin{itemize}
\item The mutual information (information transfer) is equal to the output entropy.
\item It can be shown that $H(Y|X) = 0$ ( If $x_i$ is the input, there is certainty about the output). Also $I(X;Y) = H(Y)$.
\item  The channel capacity per symbol is

\[ C_s = \mbox{ max }_{P(x_i)} H(Y) = \mbox{log}_2n \]
where $n$ is the number of symbols in $Y$.
\end{itemize}

%---------------------------------------------------------------------------------Page 252 A -%

\noindent \textbf{Capacities of special channels}
 \textbf{\emph{Noiseless Channel}}:
\begin{itemize}
\item Since a noiseless channel is both lossless and deterministic , we can say that $I(X;Y) = H(X) = H(Y)$.
The mutual information (information transfer) is equal to the output entropy). \item The channel capacity per symbol is
\[ C_s = \mbox{log}_2m = \mbox{log}_2n \]
\end{itemize}


%---------------------------------------------------------------------------------Page 252 A -%

\noindent \textbf{Capacities of special channels}
 \textbf{\emph{Binary Symmetric Channel}}:
\begin{itemize}
\item It can be shown that, for a binary symmetric channel, the the channel capacity per symbol is
\[ C_s = 1 + p\mbox{log}_2p  + 1-p\mbox{log}_2 (1-p)  \]
\end{itemize}



\noindent \textbf{Source Coding}
\begin{itemize}
\item A conversion of the output of a DMS into a sequence of binary symbols (binary code word) is
called \textbf{\emph{source coding}}.
\item  The device that performs this conversion is called the source encoder.
\item
An objective of source coding is to minimize the average bit rate required for representation of the
source by reducing the redundancy of the information source.
\end{itemize}



\noindent \textbf{Source Coding : Code Length and Code Efficiency}
\begin{itemize}

\item
Let X be a DMS with finite entropy $H(X)$ and an alphabet $\{x_1 , \ldots,  x_m\}$, each with corresponding
probabilities of occurrence $P(x_i)$. \item Let the binary code word assigned to symbol $x_i$ by
the encoder have length $n_i$ b. \item The length of a code word is the number of binary digits
in the code word. The average code word length L, per source symbol is given by

\[ E(L) = \sum ^{m}_{i=1} P(x_i) n_i \]
\end{itemize}




\noindent \textbf{Source Coding : Code efficiency and Code redundancy}
% Pg 253/254
\begin{itemize}
\item The parameter $L$ (estimated by $E(L)$) represents the average number of bits per source symbol used in the source coding process.\item
The code efficiency is defined as \[\eta = {L_{min} \over L} \]where $L_{min}$ is the minimum possible value of $L$. When $\eta$ approaches unity, the codes is said to be efficient.
\item The code redundancy $\gamma$ is defined as $\gamma = 1- \eta$.
\end{itemize}



%---------------------------------------------------------------------------------------------------------------------------------------%

%page 254
\noindent \textbf{Source Coding Theorem}
\begin{itemize}
\item The source coding theorem states that for a DMS X with entropy $H(X)$, the average code word length $L$ per symbol is bounded as $L \geq H(X)$

\item Furthermore L can be made as close to H(X) as required for some suitably chosen code.
\item Thus, with $ L_{min} \geq H(X)$, the code efficiency can be rewritten as
\[\eta = {H(X) \over L} \]
\item We will use this definition for efficiency. (Remark $L$ is estimable by $E(L)$.)
\end{itemize}

%-----------------------------------------------------------------------------------------------------------------------------------------------------------%

\noindent \textbf{ ENTROPY CODING}
The design of a variable-length code such that its average code word length approaches the
entropy of the DMS is often referred to as enlmpy coding. In this section we present two examples of
entropy coding.
\begin{itemize}
\item Shannon- Fano Coiding
\item Huffman Coding
\end{itemize}

%-----------------------------------------------------------------------------------------------------------------------------------------------------------%

% Page 255
{A. Shannon-Fun Coding:}
An efficient code can be obtained by the following simple procedure, known as
Shannon- Fano algorithm:
\begin{itemize}
\item[1.] List the source symbols in order of decreasing probability.
\item[2.] Partition the set into two sets that are as close to equiprobable as possible, and assign 0 to the
upper set and 1 to the lower set.
\end{itemize}


%-----------------------------------------------------------------------------------------------------------------------------------------------------------%
% Page 256 Bottom
\noindent \textbf{B. Huffman Encoding:}
In general, Huffman encoding results in an optimum code. Thus, it is the code that has the highest
efliciency.\\ The Huffman encoding procedure is as follows:
\begin{itemize}\item[1.] List the source symbols in order of decreasing probability.
\item[2.] Combine the probabilities of the two symbols having the lowest probabilities, and reorder
the resultant probabilities; this step is called reduction 1. The same procedure is repeated until
there are two ordered probabilities remaining.
\item[3.] Start encoding with the last reduction, which consists of exactly two ordered probabilities. Assign
0 as the first digit in the code words for all the source symbols associated with the first probability;
assign 1 to the second probability.
\item[4.] Now go back and assign 0 and 1 to the second digit for the two probabilities that were combined
in the previous reduction step, retaining all assignments made in Step 3.
\item[5.] Keep regressing this way until the first column is reached.
\end{itemize}


%An example of Huffman encoding is shown in Table 10-3.
%H(X) = 2.36b/symbol
%L = 2.38 b/symbol
%\nu = 0.99

%----------------------------------------------------------------------------------%

{
Huffman coding is an entropy encoding algorithm used for lossless data compression.


}
%----------------------------------------------------------------------------------%


{
\noindent \textbf{Huffman Coding}
Huffman coding is a method of lossless data compression, and a form of entropy encoding. The basic idea is to map an alphabet to a representation for that alphabet, composed of strings of variable size, so that symbols that have a higher probability of occurring have a smaller representation than those that occur less often.

}
%----------------------------------------------------------------------------------%

{
\noindent \textbf{Huffman Coding}
The key to Huffman coding is Huffman's algorithm, which constructs an extended binary tree of minimum weighted path length from a list of weights. For this problem, our list of weights consists of the probabilities of symbol occurrence. From this tree (which we will call a Huffman tree for convenience), the mapping to our variable-sized representations can be defined.
}
%-------------------------------------------------------------------------------------------%
{
\noindent \textbf{Huffman Coding}
The mapping is obtained by the path from the root of the Huffman tree to the leaf associated with a symbol's weight. The method can be arbitrary, but typically a value of 0 is associated with an edge to any left child and a value of 1 with an edge to any right child (or vice-versa). By concatenating the labels associated with the edges that make up the path from the root to a leaf, we get a binary string. Thus the mapping is defined.
}
%-------------------------------------------------------------------------------------------%
{
\noindent \textbf{Inverse Mapping}
\begin{itemize}
\item In order to recover the symbols that make up a string from its representation after encoding, an inverse mapping must be possible. It is important that this mapping is unambiguous. \item We can show that all possible strings formed by concatenating any number of path labels in a Huffman tree are indeed unambiguous, due to the fact that it is a complete binary tree. \item That is, given a string composed of Huffman codes, there is exactly one possible way to decompose it into the individual codes.
\end{itemize}
}
%-------------------------------------------------------------------------------------------%
{

\noindent \textbf{Ambiguity}

Ambiguity occurs if there is any path to some symbol whose label is a prefix of the label of a path to some other symbol. In the Huffman tree, every symbol is a \textbf{\emph{leaf}}. Thus it is impossible for the label of a path to a leaf to be a prefix of any other path label, and so the mapping defined by Huffman coding has an inverse and decoding is possible.
}

\end{document}
















%-----------------------------------------------------------------------------------------------%
 % ULCIS
\noindent \textbf{Self Information}Self-information
This is defined by the following mathematical formula:$I(A) = −logb P(A)$

The self-information of an event measures the amount of one's surprise
evoked by the event. The negative logarithm $−logb P(A)$ can also be written as \[
log_b  {1 \over P(A)} \]
Note that log(1) = 0, and that $| − log(P(A))|$ increases as P(A) decreases
from 1 to 0. This supports our intuition from daily experience. For example,
a low-probability event tends to cause more ``surprise".



%-------------------------------------------------------------------------------------------%
{
\noindent \textbf{Example}
For a simple example, we will take a short phrase and derive our probabilities from a frequency count of letters within that phrase. The resulting encoding should be good for compressing this phrase, but of course will be inappropriate for other phrases with a different letter distribution.

"All you base are belong to us"
}


%----------------------------------------------------------------------------------%
{
\noindent \textbf{Entropy}
\begin{itemize}
\item Entropy is the uncertainty of a single random variable. \item We can define \textbf{\emph{conditional entropy }}$H(X|Y)$, which is the entropy of a random variable
conditional on the knowledge of another random variable. \item The reduction in uncertainty due to another random variable is called the \textbf{\emph{mutual information}}.
\end{itemize}
}
%----------------------------------------------------------------------------------%


{
\noindent \textbf{What is Information?}
\begin{itemize} \item Once we agree to define the information of an event ain terms of P(a), the properties (2) and (3) will be satisfied if the information in ais defined as
\[ I(a) = -log P(a)\]

\item Remark : The base of the logarithm depends on the unit of information to be used.
\end{itemize}
}



%-------------------------------------------------------------------------------------------%
{
Ambiguity occurs if there is any path to some symbol whose label is a prefix of the label of a path to some other symbol. In the Huffman tree, every symbol is a \textbf{\emph{leaf}}. Thus it is impossible for the label of a path to a leaf to be a prefix of any other path label, and so the mapping defined by Huffman coding has an inverse and decoding is possible.
}
%-------------------------------------------------------------------------------------------%
{
\noindent \textbf{Example}
For a simple example, we will take a short phrase and derive our probabilities from a frequency count of letters within that phrase. The resulting encoding should be good for compressing this phrase, but of course will be inappropriate for other phrases with a different letter distribution.

"All you base are belong to us"
}
