\documentclass[a4]{beamer}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{newlfont}
\usepackage{amsmath,amsthm,amsfonts}
%\usepackage{beamerthemesplit}
\usepackage{pgf,pgfarrows,pgfnodes,pgfautomata,pgfheaps,pgfshade}
\usepackage{mathptmx} % Font Family
\usepackage{helvet} % Font Family
\usepackage{color}
\mode<presentation> {
\usetheme{Default} % was Frankfurt
\useinnertheme{rounded}
\useoutertheme{infolines}
\usefonttheme{serif}
%\usecolortheme{wolverine}
% \usecolortheme{rose}
\usefonttheme{structurebold}
}
\setbeamercovered{dynamic}
\title[MA4413]{Statistics for Computing \\ {\normalsize MA4413 Lecture 12A}}
\author[Kevin O'Brien]{Kevin O'Brien \\ {\scriptsize kevin.obrien@ul.ie}}
\date{Autumn 2011}
\institute[Maths \& Stats]{Dept. of Mathematics \& Statistics, \\ University \textit{of} Limerick}
\renewcommand{\arraystretch}{1.5}
%------------------------------------------------------------------------%

\begin{document}
\medskip
\titlepage

%---------------------------------------------------------------------------------------------------------------------------------------%
\medskip
\noindent \textbf{Codes}
Recall from last weeks lectures, this table below where a source of
size 4 has been encoded in binary codes with symbol 0 and 1.\\ \bigskip
% Table 10-I Binary Codes
\begin{center}
\begin{tabular}{|c| c| c| c| c| c| c|}
\hline
X& Code l& Code 2& Code 3 &Code 4& Code 5& Code 6\\\hline
$x_1$& 00& 00 &0 &0 &0 &1\\
$x_2$& 01& 01 &1 &10 &01 &01\\
$x_3$ &00 &10& 00& 110& 011 &001\\
$x_4$ &11& 11& 11& 111 &0111 &0001\\\hline
\end{tabular}
\end{center}


%---------------------------------------------------------------------------------------------------------------------------------------%
\medskip
\noindent \textbf{Code Classifications}
\begin{itemize}
\item \textbf{Code 1} This code is fixed length, but not distinct. Two symbols have the same binary representation. Due to this flaw it is no longer considered.
\item \textbf{Code 2} This code is fixed length and distinct. 
\item \textbf{Code 3} This code is not uniquely decodable. Again due to this flaw, we will no longer consider it.

\end{itemize}

%---------------------------------------------------------------------------------------------------------------------------------------%
\medskip
\noindent \textbf{Prefix-free codes}
\begin{itemize}
\item A prefix-free code is one in which no codeword is a prefix in another. \item Note that every prefix-free code is decipherable, but the converse is not true. 
    \item In code 4, none of the codewords appear as prefixes for other codewords.
\item For code 5, each code word are prefixes for the subsequent codeword.
\item Both code 4 and 5 are uniquely decodable.
\item Code 6 is prefix free and uniquely decodable.
\end{itemize}

%---------------------------------------------------------------------------------------------------------------------------------------%
\medskip
\noindent \textbf{Word Length}
\begin{itemize}
\item These codes use code lengths between 1 and 4.
\item For code 2; $n_1 = n_2 = n_3 = n_4 = 2$.
\item For code 6; $n_1 = 1$, $n_2 = 2$, $n_3 = 3$,$n_4 = 4$.
\end{itemize}

%---------------------------------------------------------------------------------------------------------------------------------------%
\medskip
\noindent \textbf{Word Length}
\begin{itemize}
\item Suppose that the symbols $\{x_1, x_2, x_3, x_4\}$ appear with the following probabilities $\{0.4, 0.3, 0.2, 0.1\}$
\item The average code word length E(L) per source symbol is given by

\[ E(L) = \sum ^{m}_{i=1} P(x_i) n_i \]
\item For each code compute $E(L)$.
\end{itemize}


%---------------------------------------------------------------------------------------------------------------------------------------%
\medskip
\noindent \textbf{Word Length}
\begin{itemize}
\item \textbf{Code 1 and 2} Codes are fixed length $ E(L) = 2 $
\item \textbf{Code 3} Recall: Code is flawed \[ E(L) = (0.4 \times 1)+(0.3 \times 1)+(0.2 \times 2)+(0.1 \times 2) = 1.3 \]
\item \textbf{Code 4 } \[ E(L) = (0.4 \times 1)+(0.3 \times 2)+(0.2 \times 3)+(0.1 \times 3) = 1.9 \]
\item \textbf{Code 5 and 6} \[ E(L) = (0.4 \times 1)+(0.3 \times 2)+(0.2 \times 3)+(0.1 \times 4) = 2 \]
\end{itemize}


%---------------------------------------------------------------------------------------------------------------------------------------%
\medskip
\noindent \textbf{Word Length : Interpretation}
\begin{itemize}
\item Code 4 would require 190 binary digits to transmit 100 symbols. The transmission would be uniquely decodable.
\item Code 3 would require 130 binary digits to transmit 100 symbols. The transmission would be not be uniquely decodable, and the intended message would be unclear.
\item For the other codes, each would require 200 digits.
\item Code 4 is seemingly the best choice.
\end{itemize}

%---------------------------------------------------------------------------------------------------------------------------------------%
\medskip
\noindent \textbf{Entropy and Efficiency}
Given that the entropy of the input source is $H(X) = 1.85$b, compute the efficiency $\eta$ for each code.
\begin{itemize}
\item \textbf{Code 1 and 2} \\$ \eta = H(X) / E(L) = [1.85 / 2] \times 100\% = 92.5\% $
\item \textbf{Code 3} \\Recall that this code is flawed $ \eta = H(X) / E(L) = [1.84 / 1.3] \times 100\% = 142\% $
\item \textbf{Code 4}\\ $ \eta = H(X) / E(L) = [1.85 / 1.9] \times 100\% = 97.2\% $
\item \textbf{Code 5 and 6}\\ $ \eta = H(X) / E(L) = [1.85 / 2] \times 100\% = 92.5\% $

\end{itemize}



\medskip
\noindent \textbf{Instantaneous Codes}
\begin{itemize} \item Recall from previous lecture \item A uniquely decodable code is called an instantaneous code if the end of any code word is
recognizable without examining subsequent code symbols. \item The instantaneous codes have the property
that no code word is a prefix of another code word.  \item Codes 2,4 and 6 are prefix-free codes, hence they are instantaneous codes.\end{itemize}






%-------------------------------------------------------------------------------------------%
\end{document} 
